\chapter{Special Cases and Restrictions}
\label{chapter:special}

Having proven that \probBook is a hard problem in the previous section,
we now turn to some special instances or variations on \probBook and
either show how they can be solved efficiently or why they remain hard.
This section is the main contribution of this thesis.

One topic we are interested in is how embeddability depends
on the connectivity of the pages. Thus, we deal with both extreme
cases with regards to connectivity in the first two sections: The pages
can either all be connected~(\myref{section:connected}), or ``maximally
disconnected'' without isolated vertices~(\myref{section:matchings}), \ie consisting of perfect disjoint matchings.

We show that the connected case---quite surprisingly---admits a solution
in linear time.
For each page we get a \PQ-tree (see \myref{def:pq}) that stands for the
possible orders of the vertices in a page embedding of this single page.
The solution intersects these sets of orders to decide embeddability. 

In contrast, we are unable to provide an efficient algorithm if the edges of each
page form perfect disjoint matchings. We only manage to prove for this
case that an embeddable graph must be bipartite. Furthermore, we derive---by hand and using a computer---positive bipartite instances and the smallest bipartite counterexamples for all numbers of pages with the
exception of three pages. For three pages we are able to get a smallest counterexample
when two of the matchings form a cycle.

Another interesting restriction we make in \myref{section:trees} is to allow the order of the 
vertices on the spine to only come from the
permutations represented by a fixed \PQ-tree. We present a quadratic time algorithm
for solving the book embedding problem with this restriction if we only allow \Q-trees.
Angelini et.\,al.~\cite{angelini11} showed that a similar restriction is important by reducing \SEFECON~(see page~\pageref{prob:sefecon}) to a book embedding problem
with 2~pages constrained by a \PT-tree.

Finally, in \myref{section:multi_spine} we give a related 
variation on the book embedding problem. We now allow multiple spines~(parallel lines) in the plane
and associate every vertex with a spine the vertex has to be drawn on. 
Additionally, we demand that edges are drawn between consecutive spines, above the
topmost spine or below the bottommost spine. We show that this variation is equivalent to
a special case of the 2-page book embedding problem where the vertex order is constrained
by a \PT-tree, \ie it is indeed related to the problem
of \myref{section:trees}.

While the results we present are mostly independent of each other and small, they
still provide significant insight into the book embedding problem and
\myref{section:trees} is a step towards solving \prob{sefe}. 

\input{connected}
\input{matchings}
\input{trees}
\input{multi_spine}